---
author: Silvana M. Pesenti
date: June 01, 2019
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# SWIM - A Package for Sensitivity Analysis

The SWIM package provides weights on simulated scenarios from a stochastic model, such that a stressed model components (random variables) fulfil given probabilistic constraints (e.g. specified values for risk measures), under the new scenario weights. Scenario weights are selected by constrained minimisation of the relative entropy to the baseline model.

## Installation

You can install the SWIM package from [GitHub](https://github.com/spesenti/SWIM) with:

``` r
install.packages("spesenti/SWIM")
```
## Scope of the SWIM package  
Implemented stresses are: 

R functions          | Stress
------------------- | -------------
stress_VaR          | VaR risk measure, a quantile
stress_VaR_ES       | VaR and ES risk measures
stress_mean         | means
stress_mean_sd      | means and standard deviations
stress_mean_moment  | moments, functions of moments
stress_prob         | probabilities of intervals
stress_user         | user defined scenario weights


## Example - Stressing the VaR of a portfolio

Consider a portfolio Y = X1 + X2 + X3 + X4 + X5, where X1, X2, X3, X4, X5) are correlated normally istributed with equal mean and different standard deviations.
We stress the VaR (quantile) of the portfolio loss Y at levels 0.75 and 0.9 with an increase of 10\%.

```{r example}
set.seed(0)
SD <- round(runif(5, 30, 80))
Corr <- matrix(rep(0.5, 5^2), nrow = 5) + diag(rep(1 - 0.5, 5))
if (!requireNamespace("mvtnorm", quietly = TRUE))
   stop("Package \"mvtnorm\" needed for this function 
   to work. Please install it.")
x <- mvtnorm::rmvnorm(10^5, 
   mean =  rep(100, 5), 
   sigma = (SD %*% t(SD)) * Corr)
data <- data.frame(rowSums(x), x)
names(data) <- c("Y", "X1", "X2", "X3", "X4", "X5")
rev.stress <- stress(type = "VaR", x = data, 
   alpha = c(0.75, 0.9), q_ratio = 1.1, k = 1)
```

Summary statistics of the baseline and the stressed model can be obtained via the summary method.  

```{r summary}
summary(rev.stress, base = TRUE)
```

Visual display of the change of empirical distirbution functions of the portfolio loss Y from the baseline to the two stressed models. 

```{r plot-cdf}
plot_cdf(object = rev.stress, xCol = , base = TRUE)
```


### Sensitivity and importance rank of portfolio components

Sensitivity measures allow to assess the importance of the input components. Implemented sensitivity measures are the Kolmogorov, the Wasserstein distance and *Gamma*. *Gamma*, the *Reverse Sensitivity Measure*, defined for a random variable Z and scenario weights w by

*Gamma* = ( E(Y * w) - E(Y) / c,

where c is a normalisation constant such that |*Gamma*| <= 1, see https://doi.org/10.1016/j.ejor.2018.10.003. Loosely speaking, the Reverse Sensitivity Measure is the normalised difference between the first moment of the stressed and the baseline distributions of Y.

```{r sensitivity}
sensitivity(rev.stress, type = "all")
plot_sensitivity(rev.stress, xCol = 2:6, type = "Gamma") 
```

Ranking the input components according to the chosen sensitivity measure, in this example using *Gamma*.

```{r importance rank}
importance_rank(rev.stress, xCol = 2:6, type = "Gamma")
```

Visual display of the change of empirical distirbution functions of X_1, the portfolio component with the largest sensitivity.. 

```{r plot-cdf input}
plot_cdf(object = rev.stress, xCol = "X5", base = TRUE)
```







